Sending a Balloon to the Edge of Space

### 2018-11-14 paper/paper.rmd
edits

---
title: "Sending a Balloon to the Edge of Space"
author: Luker Bowsher, John Kim, Vivian Liu, Simon Oros, Caroline Pang, Alec Vercruysse
date: "November 13, 2018"
output: 
  pdf_document:
    number_sections: true
    keep_tex: yes
    includes:
      in_header: "preamble.tex"
fontsize: "12pt"
indent: true
abstract: This paper details 6thsense, a mission to send a weather balloon to to edge of space in order to study Earth's atmosphere. The weather balloon is outfitted with a payload that includes a camera, inside and outside temperature sensors, a barometer, multiple GPS tracking devices, as well as a host of individual experiments. The balloon most likely reached an altitude of 20,000 meters, into the stratosphere and ozone layer. The balloon was successfully recovered after approximately one and one half hours of flight.
---

@@

### Vivian Liu: Light Intensity and Solar Power
Light intensity, specifically for the purposes of solar energy, is an important part of the initiative for cleaner energy. It is also interesting to learn about how light intensity increases or decreases as it passes through the ozone layer.

@@

@@

`r kable(errors, caption="Percent errors for the group sensors used relative to standards")`

@@

>>>>

---
title: "Sending a Balloon to the Edge of Space"
author: Luke Bowsher, John Kim, Vivian Liu, Simon Oros, Caroline Pang, Alec Vercruysse
date: "November 13, 2018"
output: 
  pdf_document:
    number_sections: true
    keep_tex: yes
    includes:
      in_header: "preamble.tex"
fontsize: "12pt"
indent: true
abstract: This paper details 6thsense, a mission to send a weather balloon to to edge of space in order to study Earth's atmosphere. The weather balloon is outfitted with a payload that includes a camera, inside and outside temperature sensors, a barometer, multiple GPS tracking devices, as well as a host of individual experiments. The balloon most likely reached an altitude of 20,000 meters, into the stratosphere and ozone layer. The balloon was successfully recovered after approximately one and one half hours of flight.
---

@@

### Vivian Liu: Light Intensity and Solar Power

Light intensity, specifically for the purposes of solar energy, is an important part of the initiative for cleaner energy. It is also interesting to learn about how light intensity increases or decreases as it passes through the ozone layer.

@@

### Luke Bowsher: Acceleromter

As altitude rises, it has been previously observed that wind speed  increases. To measure how the change in wind speed affected the turbulence of the balloon, an accelerometer chip was placed inside the payload to quantify the acceleration along all three axes.

@@

Tables 5 through 9 show the calculations of the root mean squared error (RMSE), or standard error, which are each used to calculate the total errors.

```{r loading, include=F}
presse <- read.csv("data/table/presse.csv")
humse <- read.csv("data/table/humse.csv")
tempse <- read.csv("data/table/tempse.csv")
thermse <- read.csv("data/table/thermse.csv")
total_se <- read.csv("data/table/total.csv")
```

`r kable(presse, caption="Calculated Pressure Standard Error (kPa)")`

$error_{total} = \sqrt{(error_{standard})^2 + (error_{sensor})^2} = \sqrt{(error_{standard})^2 + (slope_{conversion}/2)^2} = \sqrt{(0.3441374353)^2 + (0.2554 / 2)^2} = 0.367066567 \approx 0.367$

`r kable(humse, caption="Calculated Humidity Standard Error (°C)")`

$error_{total} = \sqrt{(error_{standard})^2 + (error_{sensor})^2} = \sqrt{(error_{standard})^2 + (slope_{conversion}/2)^2} = \sqrt{(12.3143298)^2 + (0.9337 / 2)^2} = 12.323176025 \approx 12.32$

`r kable(tempse, caption="Calculated Temperature Standard Error (°C)")`

$error_{total} = \sqrt{(error_{standard})^2 + (error_{sensor})^2} = \sqrt{(error_{standard})^2 + (slope_{conversion}/2)^2} = \sqrt{(0.7755410298)^2 + (0.9747 / 2)^2} = 0.915955191 \approx 0.916$

`r kable(thermse, caption="Calculated Thermistor Standard Error (°C)")`

$error_{total} = \sqrt{(error_{standard})^2 + (error_{sensor})^2} = \sqrt{(error_{standard})^2 + (slope_{conversion}/2)^2} = \sqrt{(8.493267690)^2 + (0.657 / 2)^2} = 8.499618127 \approx 8.50$

`r kable(total_se, caption="Total Standard Errors")`

<!--`r kable(errors, caption="Percent errors for the group sensors used relative to standards")`-->

@@

## Modeled vs Actual Trajectory

```{r kml, echo=F, out.width="400px"}
knitr::include_graphics("assets/kml.png")
```
Figure 16

Figure 16 an overlay of the simulated balloon trajectory (taken 12 hours before the launch) and the actual recorded balloon trajectory. The trajectory in the back is the actual data, and the trajectory in front is the simulated. There is a gap in the actual trajectory, which is due to the fact that the logger stopped recording data midway through its flight. In addition, because the balloon popped earlier than anticipated, it was not able to gain the expected altitude of 33km so it landed closer to the launch site. However, had conditions been optimal and the balloon reached its optimal altitude, the simulation probably would have been a good predictor of actual balloon trajectory.

### 2018-11-14 paper/paper.tex
edits

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Sending a Balloon to the Edge of Space},
            pdfauthor={Luker Bowsher, John Kim, Vivian Liu, Simon Oros, Caroline Pang, Alec Vercruysse},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\usepackage[normalem]{ulem}
% avoid problems with \sout in headers with hyperref:
\pdfstringdefDisableCommands{\renewcommand{\sout}{}}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Sending a Balloon to the Edge of Space}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Luker Bowsher, John Kim, Vivian Liu, Simon Oros, Caroline Pang, Alec
Vercruysse}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{November 13, 2018}

@@

@@

\begin{longtable}[]{@{}lr@{}}
\caption{Percent errors for the group sensors used relative to
standards}\tabularnewline
\toprule
Experiments & Percent error (\%)\tabularnewline
\midrule
\endfirsthead
\toprule
Experiments & Percent error (\%)\tabularnewline
\midrule
\endhead
Pressure & 0.459\tabularnewline
Humidity & 0.659\tabularnewline
Temperature (calibrated to Fieldpiece) & 0.429\tabularnewline
Temperature (calibrated to Extech) & 1.459\tabularnewline
\bottomrule
\end{longtable}

@@

>>>>

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Sending a Balloon to the Edge of Space},
            pdfauthor={Luke Bowsher, John Kim, Vivian Liu, Simon Oros, Caroline Pang, Alec Vercruysse},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\usepackage[normalem]{ulem}
% avoid problems with \sout in headers with hyperref:
\pdfstringdefDisableCommands{\renewcommand{\sout}{}}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Sending a Balloon to the Edge of Space}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Luke Bowsher, John Kim, Vivian Liu, Simon Oros, Caroline Pang, Alec
Vercruysse}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{November 13, 2018}

@@

\subsubsection{Luke Bowsher:
Acceleromter}\label{luke-bowsher-acceleromter}

As altitude rises, it has been previously observed that wind speed
~increases. To measure how the change in wind speed affected the
turbulence of the balloon, an accelerometer chip was placed inside the
payload to quantify the acceleration along all three axes.

@@

Tables 5 through 9 show the calculations of the root mean squared error
(RMSE), or standard error, which are each used to calculate the total
errors.

\begin{longtable}[]{@{}lrll@{}}
\caption{Calculated Pressure Standard Error (kPa)}\tabularnewline
\toprule
Vernier.Measurement & Vernier.Predicted & Extech.Measured &
Square.errors\tabularnewline
\midrule
\endfirsthead
\toprule
Vernier.Measurement & Vernier.Predicted & Extech.Measured &
Square.errors\tabularnewline
\midrule
\endhead
105.77 & 1.7200 & 1.48 & 0.057\tabularnewline
149.44 & 12.8700 & 12.75 & 0.014\tabularnewline
207.92 & 27.8100 & 27.79 & 0.000\tabularnewline
274.45 & 44.8000 & 45.04 & 0.059\tabularnewline
344.59 & 62.7100 & 62.95 & 0.058\tabularnewline
409.80 & 79.3700 & 79.86 & 0.239\tabularnewline
497.34 & 101.7300 & 101.09 & 0.401\tabularnewline
RMSE & 0.3441 & \ldots{} & \ldots{}\tabularnewline
\bottomrule
\end{longtable}

\(error_{total} = \sqrt{(error_{standard})^2 + (error_{sensor})^2} = \sqrt{(error_{standard})^2 + (slope_{conversion}/2)^2} = \sqrt{(0.3441374353)^2 + (0.2554 / 2)^2} = 0.367066567 \approx 0.367\)

\begin{longtable}[]{@{}lll@{}}
\caption{Calculated Humidity Standard Error (°C)}\tabularnewline
\toprule
BME280.Predicted & Vernier.Measured & Square.error\tabularnewline
\midrule
\endfirsthead
\toprule
BME280.Predicted & Vernier.Measured & Square.error\tabularnewline
\midrule
\endhead
32.11 & 22.03 & 101.61\tabularnewline
32.12 & 21.74 & 107.74\tabularnewline
32.12 & 21.59 & 110.88\tabularnewline
\ldots{} & \ldots{} & \ldots{}\tabularnewline
81.15 & 65.87 & 233.48\tabularnewline
81.20 & 66.31 & 221.71\tabularnewline
81.25 & 66.31 & 223.20\tabularnewline
RMSE / standard error & 12.3143298 & \ldots{}\tabularnewline
\bottomrule
\end{longtable}

\(error_{total} = \sqrt{(error_{standard})^2 + (error_{sensor})^2} = \sqrt{(error_{standard})^2 + (slope_{conversion}/2)^2} = \sqrt{(12.3143298)^2 + (0.9337 / 2)^2} = 12.323176025 \approx 12.32\)

\begin{longtable}[]{@{}lll@{}}
\caption{Calculated Temperature Standard Error (°C)}\tabularnewline
\toprule
BME280.Predicted & Fieldpiece.Measured & Square.errors\tabularnewline
\midrule
\endfirsthead
\toprule
BME280.Predicted & Fieldpiece.Measured & Square.errors\tabularnewline
\midrule
\endhead
-21.05 & -20.90 & 0.023\tabularnewline
-21.16 & -21.00 & 0.026\tabularnewline
-21.20 & -21.10 & 0.010\tabularnewline
\ldots{} & \ldots{} & \ldots{}\tabularnewline
4.12 & 4.70 & 0.336\tabularnewline
4.12 & 4.60 & 0.230\tabularnewline
4.13 & 4.40 & 0.073\tabularnewline
RMSE / standard error & 0.7755410298 & \ldots{}\tabularnewline
\bottomrule
\end{longtable}

\(error_{total} = \sqrt{(error_{standard})^2 + (error_{sensor})^2} = \sqrt{(error_{standard})^2 + (slope_{conversion}/2)^2} = \sqrt{(0.7755410298)^2 + (0.9747 / 2)^2} = 0.915955191 \approx 0.916\)

\begin{longtable}[]{@{}lll@{}}
\caption{Calculated Thermistor Standard Error (°C)}\tabularnewline
\toprule
Thermistor.Predicted & Extech.Measured & Square.errors\tabularnewline
\midrule
\endfirsthead
\toprule
Thermistor.Predicted & Extech.Measured & Square.errors\tabularnewline
\midrule
\endhead
18.38 & 20.80 & 5.86\tabularnewline
18.38 & 20.70 & 5.38\tabularnewline
18.38 & 20.70 & 5.38\tabularnewline
\ldots{} & \ldots{} & \ldots{}\tabularnewline
-14.67 & -3.30 & 129.28\tabularnewline
-14.83 & -3.50 & 128.37\tabularnewline
-14.67 & -3.70 & 120.34\tabularnewline
RMSE / standard error & 8.493267690 & \ldots{}\tabularnewline
\bottomrule
\end{longtable}

\(error_{total} = \sqrt{(error_{standard})^2 + (error_{sensor})^2} = \sqrt{(error_{standard})^2 + (slope_{conversion}/2)^2} = \sqrt{(8.493267690)^2 + (0.657 / 2)^2} = 8.499618127 \approx 8.50\)

\begin{longtable}[]{@{}lr@{}}
\caption{Total Standard Errors}\tabularnewline
\toprule
Experiment\ldots{}sensor & Total.error\tabularnewline
\midrule
\endfirsthead
\toprule
Experiment\ldots{}sensor & Total.error\tabularnewline
\midrule
\endhead
Pressure & 0.367\tabularnewline
Humidity & 12.320\tabularnewline
Temperature (BME280) & 0.916\tabularnewline
Temperature (thermistor) & 8.500\tabularnewline
\bottomrule
\end{longtable}

@@

\subsection{Modeled vs Actual
Trajectory}\label{modeled-vs-actual-trajectory}

\begin{center}\includegraphics[width=400px]{assets/kml} \end{center}

Figure 16

Figure 16 an overlay of the simulated balloon trajectory (taken 12 hours
before the launch) and the actual recorded balloon trajectory. The
trajectory in the back is the actual data, and the trajectory in front
is the simulated. There is a gap in the actual trajectory, which is due
to the fact that the logger stopped recording data midway through its
flight. In addition, because the balloon popped earlier than
anticipated, it was not able to gain the expected altitude of 33km so it
landed closer to the launch site. However, had conditions been optimal
and the balloon reached its optimal altitude, the simulation probably
would have been a good predictor of actual balloon trajectory.
